\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-06-04 doc/paper.tex
final suggestions from Dominik on section 6

The RGEs for the model parameters together with the boundary
conditions form a boundary value problem.  The default two-scale RG
solver tries to find a set of model parameters consistent with all
constraints at all scales.  It does so by running iteratively between
the scales of all boundary conditions, imposing the constraints and
checking for convergence after each iteration.  This approach is
decribed in \cite{Barger:1993gh} for the MSSM and is widely
implemented in SUSY spectrum generators.

@@

      v_d(M_Z) &= \frac{2 M_Z(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \cos\beta(M_Z)}, \\
      v_u(M_Z) &= \frac{2 M_Z(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \sin\beta(M_Z)}.

@@

      A^f(M_X) &= A_0 y_f(M_X) & &(f=u,d,e),\\

@@

SUSY-scale and the pole-mass spectrum is calculated.  If the user has
chosen a specific output scale for the running \DRbar\ model
parameters by setting entry $12$ in block \code{MODSEL} in the SLHA
input file, all model parameters are run to the defined output scale.

@@

\subsection{Pole masses}
\label{sec:PoleMasses}
After the solver routine has finished and convergence has been
achieved, all $\overline{DR}$ parameters consistent with the
one-loop EWSB conditions, low energy data and all user supplied
boundary conditions are known at any scale between \code{LowScale} and
\code{HighScale}.

The physical (pole) mass spectrum can now be calculated.  \fs uses the
full one-loop self-energies and tree-level mass matrices to calculate
the pole masses at the one-loop level, which means finding the values
$p$ that solve the equation
%
\begin{align}
  0 = \det\left[p^2\unity - m^2_{f,1L}(p^2)\right],
\end{align}
%
where the one-loop mass matrices $m_{f,1L}(p^2)$ are given in terms of
the tree-level mass matrices $m_f$ and the self-energies
$\Sigma_f(p^2)$ as
%
\begin{align}
  &\text{scalars } \phi: &
  m^2_{\phi,1L}(p^2) &= m^2_{\phi} - \Sigma_\phi(p^2), \\
  &\text{majorana fermions } \chi: &
  m_{\chi,1L}(p^2) &= m_{\chi} - \frac{1}{2}\Big[
    \Sigma_\chi^S(p^2) + \Sigma_\chi^{S,T}(p^2)
    + \Big( \Sigma_\chi^{L,T}(p^2) + \Sigma_\chi^R(p^2) \Big) m_{\chi} \notag \\
    &&&\phantom{= m_{\chi} - \frac{1}{2}\Big[}
    + m_{\chi} \Big( \Sigma_\chi^L(p^2) + \Sigma_\chi^{R,T}(p^2) \Big)
  \Big], \\
  &\text{dirac fermions } \psi: &
  m_{\psi,1L}(p^2) &= m_{\psi}
  - \Sigma_\psi^S(p^2)
  - \Sigma_\psi^R(p^2) m_{\psi}
  - m_{\psi} \Sigma_\psi^L(p^2) .
\end{align}
%
Since the one-loop mass matrices depend on $p$, an iterative procedure
must be used.  \fs provides three different methods with different
precision and execution speed to determine the mass eigenvalues.
These methods are set in the model file and work as follows:
%
\begin{itemize}
\item \code{LowPoleMassPrecision}: This option provides the
  lowest precision but is also the fastest one.  Here the one-loop
  mass matrix $m_{f,1L}^\text{low}$ is calculated exactly once as
%
  \begin{align}
    \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i}
    m_{f_j}))_{ij} ,
  \end{align}
%
  where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
  matrix $m_f$.  Afterwards $m_{f,1L}^\text{low}$ is diagonalized and
  the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.

\item \code{MediumPoleMassPrecision} (default): This option
  provides calculation with medium precision with a medium execution
  time.  Here the one-loop mass matrix $m_{f,1L}^\text{medium}$ is
  calculated $n$ times as
%
  \begin{align}
    (m_{f,1L}^\text{medium})_{ij}^{(k)} = (m_{f,1L}(p^2 =
    m_{f_k}^2))_{ij} , \qquad k = 1,\ldots,n ,
  \end{align}
%
  where $m_{f_k}$ is the $k$th mass eigenvalue of the tree-level mass
  matrix $m_f$.  Afterwards, each mass matrix
  $(m_{f,1L}^\text{medium})^{(k)}$ is diagonalized and the $k$th
  eigenvalue is interpreted as pole mass $m_{f_k}^\pole$.

\item \code{HighPoleMassPrecision}: This option provides
  diagonalization with highest precision, but has also the highest
  execution time.  Here the one-loop mass matrix
  $m_{f,1L}^\text{high}$ is diagonalized $n$ times, as in the case of
  \code{MediumPoleMassPrecision}, resulting in $n$ pole masses
  $m_{f_k}^\pole$ ($k = 1,\ldots,n$).  Afterwards, the diagonalization
  is repeated, this time using the calculated pole masses
  $m_{f_k}^\pole$ for the momentum calculation $p^2 =
  (m_{f_k}^\pole)^2$.  The iteration stops until convergence is
  reached.
\end{itemize}

If higher accuracy is required additional routines with higher order
corrections can be added by setting \code{UseHiggs2LoopMSSM = True;}
in the model file. For example in the MSSM by default \fs adds
two-loop Higgs FORTRAN routines supplied by P.~Slavich from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
to add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.

Something similar is done for the NMSSM, but in this case the NMSSM
$\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}, while for
$\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces
are used though it should be understood that these are not complete in
the NMSSM. For other models since the Higgs mass is a very important
measurement and the two loop corrections can be larger than the
current experimental error \cite{Degrassi:2009yq} we recommend that
the leading log two loop corrections are estimated, by generalising
those of the MSSM or NMSSM, as has been done, for example, in the
E$_6$SSM\cite{King:2005jy}.

>>>>

As explained at the beginning of \secref{sec:SpecGenStruct}, the RGEs
plus the user-defined boundary conditions on the model parameters form
a boundary value problem.  \fs provides a default two-scale boundary
value problem solver, which tries to find a set of model parameters
consistent with all constraints at all scales.  It does so by running
iteratively between the scales of all boundary conditions, imposing
the constraints (by calling the \code{apply()} function) and checking
for convergence after each iteration.  This approach is decribed in
\cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY
spectrum generators.

@@

      v_d(M_Z) &= \frac{2 M_Z^\text{\DRbar}(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \cos\beta(M_Z)}, \\
      v_u(M_Z) &= \frac{2 M_Z^\text{\DRbar}(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \sin\beta(M_Z)}.

@@

      A^f(M_X) &= A_0 & &(f=u,d,e),\\

@@

SUSY-scale and the pole-mass spectrum is calculated, see
\secref{sec:PoleMasses}.  If the user has chosen a specific output
scale for the running \DRbar\ model parameters by setting entry $12$
in block \code{MODSEL} in the SLHA input file, all model parameters
are finally run to the defined output scale.

@@

\subsection{Pole masses}
\label{sec:PoleMasses}
After the solver routine has finished and convergence has been
achieved, all \DRbar\ parameters consistent with the EWSB conditions,
low energy data and all user-supplied boundary conditions and are
known at any scale between \code{LowScale} and \code{HighScale}.

The (physical) pole mass spectrum can now be calculated.  \fs uses the
full one-loop self-energies and tree-level mass matrices obtained from
SARAH to calculate the pole masses, which means finding the values $p$
that solve the equation
%
\begin{align}
  0 = \det\left[p^2\unity - m_{f,1L}(p^2)\right].
  \label{eq:pole-mass-def}
\end{align}
%
Here the one-loop mass matrix $m_{f,1L}(p^2)$ for field $f$ is given
in terms of the tree-level mass matrix $m_f$ and the self-energy
$\Sigma_f(p^2)$ as
%
\begin{align}
  &\text{scalars } \phi: &
  m_{\phi,1L}(p^2) &= m_{\phi} - \Sigma_\phi(p^2), \\
  &\text{majorana fermions } \chi: &
  m_{\chi,1L}(p^2) &= m_{\chi} - \frac{1}{2}\Big[
    \Sigma_\chi^S(p^2) + \Sigma_\chi^{S,T}(p^2)
    + \Big( \Sigma_\chi^{L,T}(p^2) + \Sigma_\chi^R(p^2) \Big) m_{\chi} \notag \\
    &&&\phantom{= m_{\chi} - \frac{1}{2}\Big[}
    + m_{\chi} \Big( \Sigma_\chi^L(p^2) + \Sigma_\chi^{R,T}(p^2) \Big)
  \Big], \\
  &\text{dirac fermions } \psi: &
  m_{\psi,1L}(p^2) &= m_{\psi}
  - \Sigma_\psi^S(p^2)
  - \Sigma_\psi^R(p^2) m_{\psi}
  - m_{\psi} \Sigma_\psi^L(p^2) .
\end{align}
%
Eq.~\eqref{eq:pole-mass-def} can be solved by diagonalizing the
one-loop mass matrix $m_{f,1L}(p^2)$.  However, since $m_{f,1L}(p^2)$
depends on the momentum $p$, an iterative procedure must be used.
Since this iteration can be very time consuming for large field
multiplets, \fs provides two approximative procedures with a faster
run-time.  The used procedure can be set in the model file for each
field.  They work as follows:
%
\begin{itemize}
\item \code{LowPoleMassPrecision}: This option provides the
  lowest precision but is also the fastest one.  Here the one-loop
  mass matrix $m_{f,1L}^\text{low}$ is calculated exactly once as
%
  \begin{align}
    \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i}
    m_{f_j}))_{ij} ,
  \end{align}
%
  where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
  matrix $m_f$.  Afterwards, $m_{f,1L}^\text{low}$ is diagonalized and
  the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.
  This method neglects which are formally of two-loop order.  This
  method is imprecise if the self-energy corrections to the mass
  matrix are large or the tree-level mass spectrum of the multiplet is
  very split.

\item \code{MediumPoleMassPrecision} (default): This option
  provides calculation with medium precision with a medium execution
  time.  Here the one-loop mass matrix $m_{f,1L}^\text{medium}$ is
  calculated $n$ times as
%
  \begin{align}
    (m_{f,1L}^\text{medium})_{ij}^{(k)} = (m_{f,1L}(p^2 =
    m_{f_k}^2))_{ij} , \qquad k = 1,\ldots,n ,
  \end{align}
%
  where $m_{f_k}$ is the $k$th mass eigenvalue of the tree-level mass
  matrix $m_f$.  Afterwards, each mass matrix
  $(m_{f,1L}^\text{medium})^{(k)}$ is diagonalized and the $k$th
  eigenvalue is interpreted as pole mass $m_{f_k}^\pole$.  This method
  is imprecise if the self-energy corrections to the mass matrix are
  large.

\item \code{HighPoleMassPrecision}: This option solves
  Eq.~\eqref{eq:pole-mass-def} exactly by iterating over the momentum
  $p$.  It therefore provides the determination of the pole masses
  with highest precision, but has also the highest execution time.
  Here the one-loop mass matrix $m_{f,1L}^\text{high}$ is diagonalized
  $n$ times, as in the case of \code{MediumPoleMassPrecision},
  resulting in $n$ pole masses $m_{f_k}^\pole$ ($k = 1,\ldots,n$).
  Afterwards, the diagonalization is repeated, this time using the
  calculated pole masses $m_{f_k}^\pole$ for the momentum calculation
  $p^2 = (m_{f_k}^\pole)^2$.  The iteration stops if convergence is
  reached.
\end{itemize}
%
If higher accuracy is required, two-loop corrections to the
self-energies can be added by setting \code{UseHiggs2LoopMSSM = True;}
in the MSSM or \code{UseHiggs2LoopNMSSM = True;} in the NMSSM in the
model file.  This enables two-loop Higgs FORTRAN routines supplied by
P.~Slavich from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
which add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.  Something similar is
done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$
pieces come from \cite{Degrassi:2009yq}, while for $\oatq$,
$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces are used
though it should be understood that these are not complete in the
NMSSM. For other models since the Higgs mass is a very important
measurement and the two-loop corrections can be larger than the
current experimental error \cite{Degrassi:2009yq} we recommend that
the leading log two-loop corrections are estimated, by generalising
those of the MSSM or NMSSM, as has been done, for example, in the
E$_6$SSM \cite{King:2005jy}.
